---
title: 'kdensity: Kernel density estimation with a parametric start'
author: "Jonas Moss"
date: "8 February 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width='750px', dpi=200)
```

## Introduction

Kernel density estimation with a parametric start was introduced by Nils Lid Hjort and Ingrid Glad in their 1995 paper [Nonparametric Density Estimation with a Parametric Start](https://projecteuclid.org/euclid.aos/1176324627). The idea is to start out with a parametric density before you do your kernel density estimation, so that your actual kernel density estimation will be a correction to the original parametric estimate. Why is this a good idea? Because the resulting estimator will be better than a naïve kernel density estimator whenever the true density is close to your suggestion, and probably not worse if not. Read the paper for more information.

The goal of this `R` package is to make kernel density estimation with a parametric start both routine and flexible. In order to achieve this, we have decided on emulating the behaviour of the function `density` in the `stats` package (which is included in your `R` installation) to the highest extent possible.

**NB:** This package is mint out of box! At least one core feature, the choice of bandwidth, has not been implemented, and the documentation is not fully spelled out. Come back later for more! The package only supports estimates in one dimension at the moment.

## Example
The data set `sunspot.month` is described at `R` site [here](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/sunspot.month.html). This is a good example to showcase the usefulness of kernel density estimation with a parametric start, as it is supported on positive half-line. In such a case, traditional density estimators are prone to serious *boundary bias*. However, parametric starts allows us to circumvent this by using a pre-specified parametric density supported on the half-line, for instance the exponential.

```{r small plot, echo = TRUE}
plot(kdensity::kdensity(sunspot.month, start = "exponential", bw = 16), lwd = 2)

```

Lets compare this to the behaviour of `density` and `dexp`:

```{r full plot, echo = TRUE}
y = seq(0, max(sunspot.month) + 10, by = 0.5)
plot(kdensity::kdensity(sunspot.month, start = "exponential", bw = 16), lwd = 2)
lines(y, dexp(y, 1/mean(sunspot.month)), lty = 2)
lines(density(sunspot.month, kernel = "gauss", adjust = 2), lty = 3)

```

Here `density` is easily seen to be far more  affected by boundary bias than `kdensity`, and it capturs more of features of the data than the exponential distribution does.

## How to use

The function `kdensity` takes some `data`, a kernel `kernel` and a parametric start `start`. Currently the following arguments for `kernel` are supported: `gaussian`, `laplace`, `epanechnikov`, `rectangular`, `triangular`, `biweight`, `triweight`, `tricube`, `cosine`, and `optcosine`. For the `start` parameter, the following are accepted (as strings!): `uniform`, `normal`, `gamma`, `exponential`, `inverse_gaussian`, `lognormal`, `beta`, and `laplace`. Their parameters are estimated by maximum likelihood. In addition to this, you can specify the `support` parameter, which is used to find the normalizing constant.You do not have to worry about this one though, as it is automatically deduced from the choice of `start`. Still, if you want to have support on [0, 12.3], it is possible.

Finally, instead of specifying `start` as a string, you can send a `list` containing two named functions as elements: `density` is a density function with named parameters, while `estimator` is a function of `data` that estimates the parameters for you. For example, take a look at this implementation of the inverse Gaussian distribution:


```{r lnorm, echo = TRUE}
start_inverse_gaussian = list(
  density = statmod::dinvgauss,
  estimator = function(data) {
    c(mean       = mean(data),
      dispersion = mean(1/data - 1/mean(data)))
  }
)
```

Using this definition we can run a new `kdensity` function as follows:

```{r lnormplot, echo = TRUE}
plot(kdensity::kdensity(sunspot.month + 1, start = start_inverse_gaussian, support = c(0, Inf), bw = 16), lwd = 2)
```

Here `sunspot.month + 1` is used inside `kdensity` in order to make the data strictly positive.

The `plot` function works just as in the case of `stats::density`. Moreover, `lines` and `points` does as well. Since the return value of `kdensity` is a function, it is callable, as in:

```{r callable, echo = TRUE}
new_density = kdensity::kdensity(sunspot.month + 1, start = "lognormal", bw = 16)
new_density(56.7)
```

## Installation
First you need to install the package `devtools` from `CRAN`. Then you can call:
```{r install, echo = TRUE, eval = FALSE}
devtools::install_github("JonasMoss/kdensity")
```
Enjoy!
